---
title: "Final Project"
format: html
---

## Background

Stream water concentrations can be affected by watershed disturbances like hurricanes, which can be studied using long-term weekly records of stream water chemistry in the Luquillo Mountains of Puerto Rico before and after Hurricane Hugo in 1989.

Four watersheds, Puente Roto Mameyes (PRM) and Bisley Quebradas 1, 2, and 3 (BQ1, BQ2, and BQ3) are sites located within the Luquillo Experimental Forest (LEF).

This paper serves to recreate Figure 3 of the study in Schaefer et al. (2000), which represents 9-week moving averages of concentrations of various nutrients before and after the hurricane in each of the four watersheds.

## Data

### Set up and read in data
```{r}
#| warning: false
#| message: false
# Load packages
library(here)
library(tidyverse)
library(dplyr)
library(janitor)
library(patchwork)

# Source functions
source(here("R", "moving_average.R"))

# Read in csv files
PRM <- read_csv(here("data", "RioMameyesPuenteRoto.csv"))
BQ1 <- read_csv(here("data", "QuebradaCuenca1-Bisley.csv"))
BQ2 <- read_csv(here("data", "QuebradaCuenca2-Bisley.csv"))
BQ3 <- read_csv(here("data", "QuebradaCuenca3-Bisley.csv"))

```

### Clean and filter data
```{r}
#| message: false
# Join all data frames, clean up column names
full_join <- PRM %>% 
  full_join(BQ1) %>% 
  full_join(BQ2) %>% 
  full_join(BQ3) %>% 
  clean_names() 

# Filter data 
filtered_df <- full_join %>% 
  select("sample_id", "sample_date", "no3_n", "k", "mg", "ca", "nh4_n") %>%
  filter(lubridate::year(sample_date) %in% c(1988:1994)) 

# Use write_csv() to save df!
```

## Methods

```{r}
## Determine rolling averages of each nutrient

# Rolling average of no3_n concentration for all sites 
filtered_df$calc_rolling_no3_n <- sapply(
  # variable to loop over/apply to all of
  filtered_df$sample_date,
  moving_average,
  dates = filtered_df$sample_date,
  conc = filtered_df$no3_n,
  win_size_wks = 9
)

# Rolling average of k concentration for all sites 
filtered_df$calc_rolling_k <- sapply(
  # variable to loop over/apply to all of
  filtered_df$sample_date,
  moving_average,
  dates = filtered_df$sample_date,
  conc = filtered_df$k,
  win_size_wks = 9
)

# Rolling average of mg concentration for all sites 
filtered_df$calc_rolling_mg <- sapply(
  # variable to loop over/apply to all of
  filtered_df$sample_date,
  moving_average,
  dates = filtered_df$sample_date,
  conc = filtered_df$mg,
  win_size_wks = 9
)

# Rolling average of ca concentration for all sites 
filtered_df$calc_rolling_ca <- sapply(
  # variable to loop over/apply to all of
  filtered_df$sample_date,
  moving_average,
  dates = filtered_df$sample_date,
  conc = filtered_df$ca,
  win_size_wks = 9
)

# Rolling average of nh4_n concentration for all sites 
filtered_df$calc_rolling_nh4_n <- sapply(
  # variable to loop over/apply to all of
  filtered_df$sample_date,
  moving_average,
  dates = filtered_df$sample_date,
  conc = filtered_df$nh4_n,
  win_size_wks = 9
)


# Try a pivot longer data set to be able to graph by nutrient
pivot <- filtered_df %>% 
  pivot_longer(cols = c("calc_rolling_no3_n", "calc_rolling_k", "calc_rolling_mg", "calc_rolling_ca", "calc_rolling_nh4_n"),
               names_to = "calc_rolling",
               values_to = "avg")
```

## Results

### Patchwork figure
```{r}
#| warning: false
# GGplot for each nutrient

# Create k graph
figurek <- ggplot(filtered_df, aes(x = sample_date, y = calc_rolling_k)) +
  geom_line(aes(linetype = sample_id)) +
  labs(x = "Year",
       y = "K mg/l")

# Create no3_n graph
figureno3_n <- ggplot(filtered_df, aes(x = sample_date, y = calc_rolling_no3_n)) +
  geom_line(aes(linetype = sample_id)) +
  labs(x = "Year",
       y = "NO3-N ug/l")

# Create mg graph
figuremg <- ggplot(filtered_df, aes(x = sample_date, y = calc_rolling_mg)) +
  geom_line(aes(linetype = sample_id)) +
  labs(x = "Year",
       y = "Mg mg/l")

# Create ca graph
figureca <- ggplot(filtered_df, aes(x = sample_date, y = calc_rolling_ca)) +
  geom_line(aes(linetype = sample_id)) +
  labs(x = "Year",
       y = "Ca mg/l")

# Create nh4_n graph
figurenh4_n <- ggplot(filtered_df, aes(x = sample_date, y = calc_rolling_nh4_n)) +
  geom_line(aes(linetype = sample_id)) +
  labs(x = "Year",
       y = "NH4-N ug/l")

# Combine figures into a patchwork figure
patchwork <- (figurek/figureno3_n/figuremg/figureca/figurenh4_n) +
  plot_layout(axes = "collect", guides = "collect")
patchwork & theme_light() +
  theme(
    legend.title = element_text("Site")
  )
```

### Alternative: facet-wrapped figure
```{r}
#| warning: false
# Facet-wrapped figure
facet <- ggplot(pivot, aes(x = sample_date, y = avg)) +
  geom_line(aes(linetype = sample_id)) +
  facet_wrap(~calc_rolling, scales = "free_y", ncol = 1)

facet
```
